COG-ED Manuscript

### 2022-01-21 06_Paper/COG-ED/Manuscript.Rmd
put pilot results into methodsection & fixed typos

A way to quantify the subjective cost of each n-back level has been developed by Westbrook, Kester, and Braver [-@Westbrook2013], called the Cognitive Effort Discounting Paradigm (COG-ED).
First, the participants complete the n-back levels to familiarize themselves with the task.
Then, 1-back is compared with each more difficult level by asking the participants to decide between receiving 2\$ for the more difficult level or 1\$ for 1-back.
If they choose the more difficult level, the reward for 1-back increases by 0.50\$, if the choose 1-back, it decreases by 0.50\$.
This is repeated five more times, with each adjustment of the 1-back reward being half of the previous step, while the reward for the more difficult level remains fixed at 2\$.
The idea is to estimate the point of subjective equivalence, i.e. the monetary ratio at which both offers are equally preferred [@Westbrook2013].
The subjective value (SV) of each difficult level is then calculated by dividing the final reward value of 1-back by the fixed 2\$ reward.
Westbrook et al. [-@Westbrook2013] used these SVs to investigate inter-individual differences in effort discounting (ED).
Younger participants showed lower ED, i.e. they needed a lower monetary incentive for choosing the more difficult levels over 1-back.

@@

The present study changes one fundamental assumption of the original COG-ED paradigm: That the easiest n-back level has the highest SV.
We adapted the COG-ED paradigm in such a way that it allows the computation of SVs for different n-back levels without presuming that all individuals inherently prefer the easiest level.
Figure 1 illustrates how different modifications of the COG-ED paradigm return SVs that do or do not reflect the true preference of a hypothetical participant, who likes 2-back most, 3-back less, and 1-back least.
The COG-ED paradigm sets the SV of 1-back to 1, regardless of the response pattern.
Adding a comparison of 2-back and 3-back allows the SVs of those two levels to be more differentiated, but leaves the SV of 1-back unchanged.
Adding three more comparisons of the same levels but using the easier level as reference does approach the true preference, but has two disadvantages.
First, the SVs are still distorted by the SVs returned by the original paradigm, and second, having more task levels would lead to an exponential increase in comparisons.
Therefore, the solution lies in reducing the number of necessary comparisons by presenting only one ED round for each possible pair of levels, and by starting each round with a choice between equal prices.
For example, the participant is presented with the choice of receiving 1€ for 2-back or 1€ for 4-back.
The level chosen by the participant will then be used as the level with a flexible value, which starts at 1€ and is changed in every iteration.
The level that was not chosen will be set to a fixed value of 2€.
This procedure allows to compute SVs based on actual individual preference instead of objective task load.
Each level's SV is calculated as the mean of this level's SVs from all comparisons in which it appeared.
If the participant has a clear preference for one level, this level's SV will be 1.
If not, then no level's SV will be 1, but each level's SV can still be interpreted as an absolute and relative value, so each participant's ED behaviour can still be quantified.
Since we also aim to establish this paradigm for the assessment of tasks with no objective task load, e.g. emotion regulation tasks, we call it the Cognitive and Emotion Regulation Effort Discounting Paradigm (CERED).
In the present study, we will validate the CERED paradigm by conceptually replicating the findings of Westbrook et al. [-@Westbrook2013].
Additionally, we will compare the ED behaviour of participants regarding the n-back task and an emotion regulation task.
The full results of the latter will be published in a second Registered Report.
The COG-ED paradigm has been applied to tasks with different domains before, showing that SVs across task domains correlate [@Crawford2021], but these tasks had an objective order of task load, which is not the case for emotion regulation.

@@

The sample of the pilot study consisted of $N=$ `r length(setindex)-1` participants (`r round((table(data_quest$gender)[2]/(length(setindex)-1))*100, digits = 1)`% female, $M=$`r round(mean(data_quest$age, na.rm = TRUE), digits = 1)` ($SD=$`r round(sd(data_quest$age, na.rm = TRUE), digits = 1)`) years old).
One participant's data was removed because they misunderstood the instruction.
Due to a technical error the subjective task load data of one participant was incomplete, so the hypotheses involving the NASA Task Load Index were analyzed with $n=$ `r length(setindex)-2` data sets.

Healthy participants aged 18 to 30 years will be recruited using the software *ORSEE* [@Greiner2015].
Participants will fill out the personality questionnaires online and then visit the lab for two sessions one week apart.
NFC will be assessed using the 16-item short form of the Need for Cognition Scale [@Bless1994; @Cacioppo1984].
Responses to each item (e.g., "Thinking is not my idea of fun", recoded) will be recorded on a 7-point Likert scale.
The NFC scale shows comparably high internal consistency [Cronbach's $\alpha$ \> .80; @Bless1994; @Fleischhauer2010].
Several other personality questionnaires will be used in this study but are the topic of the Registered Report for the second lab session.
A full list of measures can be found in our [Github repository](https://github.com/ChScheffel/CERED).
In the first session, participants provide informed consent and demographic data before completing the computer-based paradigm.
The paradigm starts with the n-back levels one to four, presented sequentially with two runs per level, consisting of 64 consonants (16 targets, 48 non-targets) per run.
The levels are referred to by color (1-back black, 2-back red, 3-back blue, 4-back green) to avoid anchor effects in the ED procedure.
To assess perceived task load, we will use the 6-item NASA Task Load Index (NASA-TLX) [@Hart1988], where participants evaluate their subjective perception of mental load, physical load, effort, frustration, performance, and time pressure during the task on a 20-point scale.
After each level, participants fill out the NASA-TLX on a tablet.
Then, they complete the ED procedure on screen, where each possible pairing of the four n-back levels is presented in a randomized order.
Participants are instructed to decide as realistically as possible, because one of their choices from the last iteration steps will be randomly chosen for one final run of n-back.
This is only done to incentivise truthful behavior in the ED procedure, so the n-back data of this part will not be analyzed.
The second session consists of an emotion regulation task with negative pictures and the instruction to suppress facial reactions, detach cognitively from the picture content, and distract oneself, respectively.
The paradigm follows the same structure of task and ED procedure, but participants can decide which strategy they want to reapply in the last block.
Participants will receive 30€ in total or course credit for participation.
Study data will be collected and managed using REDCap electronic data capture tools hosted at Technische Universität Dresden [@Harris2009; @Harris2019].

A sample size analysis with G*Power [@Faul2007; @Faul2009], based on the results of the ANOVA of Westbrook et al. [-@Westbrook2013] which showed an increase in reaction time with higher n-back levels, indicated that we should collect data from at least 53 participants, assuming $\eta2=0.04$, $\alpha=.05$, and $\beta=.95$.
The power analyses of all other hypotheses yielded smaller necessary sample sizes.
To account for technical errors and exclusions of physiological data of the second lab session due to excessive noise, we aim to collect data of 60 to 70 participants.

Data collection and analysis will not be performed blind to the conditions of the experiments.
We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom.
The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 100 ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021].
Aggregated RT values will be described using the median and the median of absolute deviation (MAD) as robust estimates of center and variability, respectively [@Lachaud2011].
Error- and post-error trials will be excluded in repeated measures analyses of variance (rmANOVA) and controlled for in multi-level-model (MLM), because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
To test our hypotheses, we will perform a series of rmANOVAs and an MLM with orthogonal sum-to-zero contrasts in order to meaningfully interpret results [@Singmann2019]. 
Declining performance will be investigated by calculating an rmANOVA with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
An rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To ensure the validity of the MLM, we will conduct a specification curve analysis [@Simonsohn2020], which will include 63 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, 2, 2.5, or 3 $MAD$ from the median, RTs below 100 or 200 ms), and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The MLM will be run with each of the 63 pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 100 ms).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.

The association of ED and NFC will be examined with a regression using the AUC of each participant's SVs to predict their NFC score.
A second regression will additionally include the mean of the NASA-TLX subscales' AUCs of each participant as a predictor.
Since we do not have a fixed SV of 1 for 1-back, we cannot apply the "AUC" computation of Westbrook et al. [-@Westbrook2013], which was the mean of the AUCs of the SVs of each higher n-back level and 1-back, yielding values between 0 and 1.
Consequently, we will choose a different way of quantifying the individual degree of ED.
A classic AUC cannot differentiate between a subject who prefers 1-back and a subject who prefers 4-back if the magnitude of the ascent is the same, but it can reflect the overall willingness to exert effort.
This is the opposite for the sum of the ascent between SVs.
Therefore, we multiply both indicators, arriving at a value reflecting both degree and direction of preference, called $AxAUC$.

The data of this study can be downloaded from osf.io/vnj8x/.

@@

We collected data from $N = 15$ participants.
One participant's NASA-TLX data is incomplete due to a technical error, so hypotheses 1c and 3b are analyzed using the 14 complete data sets.
The results showed increases in subjective and objective task load measures with higher n-back level.
Importantly, SVs were lower for higher n-back levels, but not different between 1- and 2-back, which can be considered preliminary proof-of-concept, as this phenomon can only emerge in this version of the paradigm.
The MLM revealed that n-back level was a reliable predictor of SV, even after controlling for declining task performance (d’ and RT) as well as correct and post-correct answers, while NFC was not.
The specification curve analysis showed that this pattern was true for all 63 pipelines.
And finally, while the $AxAUC$ value did not predict any amount of variance in individual NFC scores, the AUC of NASA-TLX scores did.
All results are detailed in the *Supplementary Material*.

@@

Predictor AUC NASA-TLX:
`r apa_print(hypothesis3b)$estimate$h3_data_auc_ntlx`

>>>>

A way to quantify the subjective cost of each n-back level has been developed by Westbrook, Kester, and Braver [-@Westbrook2013], called the Cognitive Effort Discounting Paradigm (COG-ED).
First, the participants complete the n-back levels to familiarize themselves with the task.
Then, 1-back is compared with each more difficult level by asking the participants to decide between receiving 2\$ for the more difficult level or 1\$ for 1-back.
If they choose the more difficult level, the reward for 1-back increases by 0.50\$, if they choose 1-back, it decreases by 0.50\$.
This is repeated five more times, with each adjustment of the 1-back reward being half of the previous step, while the reward for the more difficult level remains fixed at 2\$.
The idea is to estimate the point of subjective equivalence, i.e. the monetary ratio at which both offers are equally preferred [@Westbrook2013].
The subjective value (SV) of each difficult level is then calculated by dividing the final reward value of 1-back by the fixed 2\$ reward.
Westbrook et al. [-@Westbrook2013] used these SVs to investigate inter-individual differences in effort discounting (ED).
Younger participants showed lower ED, i.e. they needed a lower monetary incentive for choosing the more difficult levels over 1-back.

@@

The present study changes one fundamental assumption of the original COG-ED paradigm: That the easiest n-back level has the highest SV.
We adapted the COG-ED paradigm in such a way that it allows the computation of SVs for different n-back levels without presuming that all individuals inherently prefer the easiest level.
Figure\ 1 illustrates how different modifications of the COG-ED paradigm return SVs that do or do not reflect the true preference of a hypothetical participant, who likes 2-back most, 3-back less, and 1-back least.
The COG-ED paradigm sets the SV of 1-back to 1, regardless of the response pattern.
Adding a comparison of 2-back and 3-back allows the SVs of those two levels to be more differentiated, but leaves the SV of 1-back unchanged.
Adding three more comparisons of the same levels but using the easier level as reference does approach the true preference, but has two disadvantages.
First, the SVs are still distorted by the SVs returned by the original paradigm, and second, having more task levels would lead to an exponential increase in comparisons.
Therefore, the solution lies in reducing the number of necessary comparisons by presenting only one ED round for each possible pair of levels, and by starting each round with a choice between equal prices.
For example, the participant is presented with the choice of receiving 1€ for 2-back or 1€ for 4-back.
The level chosen by the participant will then be used as the level with a flexible value, which starts at 1€ and is changed in every iteration.
The level that was not chosen will be set to a fixed value of 2€.
This procedure allows to compute SVs based on actual individual preference instead of objective task load.
Each level's SV is calculated as the mean of this level's SVs from all comparisons in which it appeared.
If the participant has a clear preference for one level, this level's SV will be 1.
If not, then no level's SV will be 1, but each level's SV can still be interpreted as an absolute and relative value, so each participant's ED behaviour can still be quantified.
Since we also aim to establish this paradigm for the assessment of tasks with no objective task load, e.g. emotion regulation tasks, we call it the Cognitive and Emotion Regulation Effort Discounting Paradigm (CERED).
In the present study, we will validate the CERED paradigm by conceptually replicating the findings of Westbrook et al. [-@Westbrook2013].
Additionally, we will compare the ED behaviour of participants regarding the n-back task and an emotion regulation task.
The full results of the latter will be published in a second Registered Report.
The COG-ED paradigm has been applied to tasks with different domains before, showing that SVs across task domains correlate [@Crawford2021], but these tasks had an objective order of task load, which is not the case for emotion regulation.

@@

@@

The sample of the pilot study consisted of $N=`r length(setindex)-1`$ participants (`r round((table(data_quest$gender)[2]/(length(setindex)-1))*100, digits = 1)`% female, $M=`r round(mean(data_quest[,"age"], na.rm = TRUE), digits = 1)`$ ($SD=`r round(sd(data_quest[,"age"], na.rm = TRUE), digits = 1)`$)\ years old).
One participant's data was removed because they misunderstood the instruction.
Due to a technical error the subjective task load data of one participant was incomplete, so the hypotheses involving the NASA Task Load Index were analyzed with $n=`r length(setindex)-2`$ data sets.
The results showed increases in subjective and objective task load measures with higher n-back level.
Importantly, SVs were lower for higher n-back levels, but not different between 1- and 2-back, which can be considered preliminary proof-of-concept, as this phenomon can only emerge in this version of the paradigm.
The MLM revealed that n-back level was a reliable predictor of SV, even after controlling for declining task performance (d’ and RT) as well as correct and post-correct answers, while NFC was not.
The specification curve analysis showed that this pattern was true for all 63 pipelines.
And finally, while the $AxAUC$ value did not predict any amount of variance in individual NFC scores, the $AUC$ of NASA-TLX scores did.
All results are detailed in the *Supplementary Material*.

Healthy participants aged 18 to 30\ years will be recruited using the software *ORSEE* [@Greiner2015].
Participants will fill out the personality questionnaires online and then visit the lab for two sessions one week apart.
NFC will be assessed using the 16-item short form of the Need for Cognition Scale [@Bless1994; @Cacioppo1984].
Responses to each item (e.g., "Thinking is not my idea of fun", recoded) will be recorded on a 7-point Likert scale.
The NFC scale shows comparably high internal consistency (Cronbach's $\alpha$ \> .80) [@Bless1994; @Fleischhauer2010].
Several other personality questionnaires will be used in this study but are the topic of the Registered Report for the second lab session.
A full list of measures can be found in our [Github repository](https://github.com/ChScheffel/CERED).
In the first session, participants provide informed consent and demographic data before completing the computer-based paradigm.
The paradigm starts with the n-back levels one to four, presented sequentially with two runs per level, consisting of 64 consonants (16 targets, 48 non-targets) per run.
The levels are referred to by color (1-back black, 2-back red, 3-back blue, 4-back green) to avoid anchor effects in the ED procedure.
To assess perceived task load, we will use the 6-item NASA Task Load Index (NASA-TLX) [@Hart1988], where participants evaluate their subjective perception of mental load, physical load, effort, frustration, performance, and time pressure during the task on a 20-point scale.
After each level, participants fill out the NASA-TLX on a tablet.
Then, they complete the ED procedure on screen, where each possible pairing of the four n-back levels is presented in a randomized order.
Participants are instructed to decide as realistically as possible, because one of their choices from the last iteration steps will be randomly chosen for one final run of n-back.
This is only done to incentivise truthful behavior in the ED procedure, so the n-back data of this part will not be analyzed.
The second session consists of an emotion regulation task with negative pictures and the instruction to suppress facial reactions, detach cognitively from the picture content, and distract oneself, respectively.
The paradigm follows the same structure of task and ED procedure, but participants can decide which strategy they want to reapply in the last block.
Participants will receive 30€ in total or course credit for participation.
Study data will be collected and managed using REDCap electronic data capture tools hosted at Technische Universität Dresden [@Harris2009; @Harris2019].

A sample size analysis with *G$\times$Power* [@Faul2007; @Faul2009], based on the results of the ANOVA of Westbrook et al. [-@Westbrook2013] which showed an increase in reaction time with higher n-back levels, indicated that we should collect data from at least 53 participants, assuming $\eta2=0.04$, $\alpha=.05$, and $\beta=.95$.
The power analyses of all other hypotheses yielded smaller necessary sample sizes.
To account for technical errors and exclusions of physiological data of the second lab session due to excessive noise, we aim to collect data of 60 to 70 participants.

Data collection and analysis will not be performed blind to the conditions of the experiments.
We aim to conduct all analysis as described in Westbrook et al. [-@Westbrook2013], but the level of detail was not always sufficient, so there might be deviations regarding data cleaning and degrees of freedom.
The performance measure *d'* will be computed as the difference of the *z*-transformed hit rate and the *z*-transformed false alarm rate [@Macmillan1990].
Reaction time (RT) data will be trimmed by excluding all trials with responses faster than 100\ ms, as the relevant cognitive processes cannot have been completed before [@Whelan2008; @Berger2021].
Aggregated RT values will be described using the median and the median of absolute deviation ($MAD$) as robust estimates of center and variability, respectively [@Lachaud2011].
Error- and post-error trials will be excluded in repeated measures analyses of variance (rmANOVA) and controlled for in multi-level-model (MLM), because RT on the latter is longer due to more cautious behavior [@Dutilh2012; @Houtman2012].
To test our hypotheses, we will perform a series of rmANOVAs and an MLM with orthogonal sum-to-zero contrasts in order to meaningfully interpret results [@Singmann2019]. 
Declining performance will be investigated by calculating an rmANOVA with three paired contrasts comparing *d'* between two levels of 2-, 3-, and 4-back at a time.
Another rmANOVA with three paired contrasts will be computed to compare the mean RT between two levels of 2-, 3-, and 4-back at a time.
To investigate changes in NASA-TLX ratings, six rmANOVAs will be computed, one for each NASA-TLX subscale, and each with six paired contrasts comparing the ratings between two levels of 1-, 2-, 3-, and 4-back at a time.
For each ED round, SVs will be calculated by adding or subtracting 0.015625 from the last monetary value of the flexible level, depending on the participant's last choice.
Then, these final monetary values will be divided by 2€, and the SV of each level per participant will be computed by averaging all final values of each level, regardless of whether it was fixed or flexible.
An rmANOVA with six paired contrasts will be computed, comparing the SVs between two levels of 1-, 2-, 3-, and 4-back at a time.
Tukey method will be used for the paired contrasts of each rmANOVA, including *p*-value adjustment.

To ensure the validity of the MLM, we will conduct a specification curve analysis [@Simonsohn2020], which will include 63 possible preprocessing pipelines of the RT data.
These pipelines specify which transformation was applied (none, log, inverse, or square-root), which outliers were excluded (none, 2, 2.5, or 3\ $MAD$ from the median, RTs below 100\ or\ 200\ ms), and across which dimensions the transformations and exclusions were applied (across/within subjects and across/within n-back levels).
The MLM will be run with each of the 63\ pipelines, which will also include our main pipeline (untransformed data, exclusion of RTs below 100\ ms).
The ratio of pipelines that lead to significant versus non-significant effects will provide an indication of how robust the effect actually is.

The association of ED and NFC will be examined with a regression using the $AUC$ of each participant's SVs to predict their NFC score.
A second regression will additionally include the mean of the NASA-TLX subscales' $AUCs$ of each participant as a predictor.
Since we do not have a fixed SV of 1 for 1-back, we cannot apply the "$AUC$" computation of Westbrook et al. [-@Westbrook2013], which was the mean of the $AUCs$ of the SVs of each higher n-back level and 1-back, yielding values between 0\ and\ 1.
Consequently, we will choose a different way of quantifying the individual degree of ED.
A classic $AUC$ cannot differentiate between a subject who prefers 1-back and a subject who prefers 4-back if the magnitude of the ascent is the same, but it can reflect the overall willingness to exert effort.
This is the opposite for the sum of the ascent between SVs.
Therefore, we multiply both indicators, arriving at a value reflecting both degree and direction of preference, called $AxAUC$.

The data of this study can be downloaded from [osf.io/vnj8x/](https://osf.io/vnj8x/).

@@

Predictor $AUC$ NASA-TLX:
`r apa_print(hypothesis3b)$estimate$h3_data_auc_ntlx`
